Faster generic IND-CCA2 secure KEM using ``encrypt-then-MAC''

### 2024-09-18 etm-transform/manuscript/main.tex
Improved KEX RTT comparison table
updated with Kyber round 3 data

%%%% 1. DOCUMENTCLASS %%%%
%
% TODO: Algo 9, 10, 11 should be one algorithms
%   similarly agorithm 13-21, 22-24, 25-30
% TODO: the implementation should be called ML-KEM+
% TODO: Table 2 should be called ML-KEM+, as well
% TODO: Table 3 should be broken up: pk/sk/keygen measurements should be one table, ct/encap/decap measurements should be one table for clearer comparison "
% TODO: Table 4-6 need double lines to separate security level
% TODO: need more thorough references
% TODO: need better intro and conclusion (but later)
% \documentclass[journal=tches,submission,notanonymous]{iacrtrans}
\documentclass[journal=tches,submission]{iacrtrans}
%%%% NOTES:
% - Change "journal=tosc" to "journal=tches" if needed
% - Change "submission" to "final" for final version
% - Add "notanonymous" to reveal authors
% - Add "spthm" for LNCS-like theorems

@@

\subsection{Choosing a message authenticator}
When instantiating \texttt{ML-KEM-ETM}, there are a variety of message authentication codes to choose from. We selected four instances covering a wide range of designs: \begin{enumerate}
    \item Poly1305, a Carter-Wegman style MAC operating on a prime field
    \item GMAC (AES-256-GCM), a Carter-Wegman style MAC operating on a binary field
    \item CMAC (AES-256-CBC), a CBC-MAC
    \item KMAC-256, a keyed hash function based on Keccak
\end{enumerate}

We tested each MAC's throughput by measuring the CPU cycles needed to compute a tag on the ciphertext returned by \texttt{K-PKE}. The length of the ciphertext varies depending on the security level. The measurements are summarized in table \ref{tbl:mac-performance}.

@@

@@

    \begin{tabular}[t]{|cc|}
        \hline
        \multicolumn{2}{|c|}{\bf ML-KEM-512} \\
        \hline
        \multicolumn{2}{|c|}{size parameters (bytes)} \\
        pubkey size & 800 \\
        seckey size & 1632 \\
        ciphertext size & 768 \\
        \hline
        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
        Median & 75945 \\
        Average & 76171 \\
        \hline
    \end{tabular}
    \begin{tabular}[t]{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant} 
        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
        \cline{2-5}
        & Median & Average & Median & Average \\
        \hline
        \texttt{ML-KEM-512} & 91467 & 92065 & 121185 & 121650 \\
        \hline
        \texttt{Kyber512} & ??? & ??? & ??? & ??? \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 93157 & 93626 & 33733 & 33908 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ GMAC & 97369 & 97766 & 37725 & 37831 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ CMAC & 99739 & 99959 & 40117 & 39943 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ KMAC & 101009 & 101313 & 40741 & 40916 \\
        \hline
    \end{tabular}\vspace{0.3cm}

    \begin{tabular}[t]{|cc|}
        \hline
        \multicolumn{2}{|c|}{\bf ML-KEM-768} \\
        \hline
        \multicolumn{2}{|c|}{size parameters (bytes)} \\
        pubkey size & 1184 \\
        seckey size & 2400 \\
        ciphertext size & 1088 \\
        \hline
        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
        Median & 129895 \\
        Average & 130650 \\
        \hline
    \end{tabular}
    \begin{tabular}[t]{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant} 
        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
        \cline{2-5}
        & Median & Average & Median & Average \\
        \hline
        \texttt{ML-KEM-768} & 136405 & 147400 & 186445 & 187529 \\
        \hline
        \texttt{Kyber768} & ??? & ??? & ??? & ??? \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 146405 & 146860 & 43315 & 43463 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ GMAC & 149525 & 150128 & 46513 & 46706 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ CMAC & 153139 & 153735 & 49841 & 50074 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ KMAC & 155219 & 155848 & 52415 & 52611 \\
        \hline
    \end{tabular}\vspace{0.3cm}

    \begin{tabular}[t]{|cc|}
        \hline
        \multicolumn{2}{|c|}{\bf ML-KEM-1024} \\
        \hline
        \multicolumn{2}{|c|}{size parameters (bytes)} \\
        pubkey size & 1568 \\
        seckey size & 3168 \\
        ciphertext size & 1568 \\
        \hline
        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
        Median & 194921 \\
        Average & 195465 \\
        \hline
    \end{tabular}
    \begin{tabular}[t]{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant} 
        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
        \cline{2-5}
        & Median & Average & Median & Average \\
        \hline
        \texttt{ML-KEM-1024} & 199185 & 199903 & 246245 & 247320 \\
        \hline
        \texttt{Kyber1024} & ??? & ??? & ??? & ??? \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 205763 & 206499 & 51375 & 51562 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 208805 & 209681 & 54573 & 54780 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 213667 & 214483 & 59175 & 59408 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 216761 & 217468 & 62269 & 62516 \\
        \hline
    \end{tabular}
\end{table}

@@

\begin{table}[H]
    \centering
    \caption{Unauthenticated key exchange RTT comparison}\label{tbl:ke-rtt}
    \begin{tabular}{|c|c|c|}
        \hline
        KEM & median RTT ($\us$) & average RTT ($\us$) \\
        \hline
        ML-KEM-512 & 92 & 97 \\
        \hline
        ML-KEM-512 + Poly1305 & 70 & 72 \\
        \hline
        ML-KEM-512 + GMAC & 73 & 76 \\
        \hline
        ML-KEM-512 + CMAC & 75 & 79 \\
        \hline
        ML-KEM-512 + KMAC & 76 & 78 \\
        \hline 
        ML-KEM-768 & 135 & 140 \\
        \hline
        ML-KEM-768 + Poly1305 & 99 & 104 \\
        \hline
        ML-KEM-768 + GMAC & 101 & 105 \\
        \hline
        ML-KEM-768 + CMAC & 103 & 109 \\
        \hline
        ML-KEM-768 + KMAC & 103 & 107 \\
        \hline
        ML-KEM-1024 & 193 & 199 \\
        \hline
        ML-KEM-1024 + Poly1305 & 138 & 141 \\
        \hline
        ML-KEM-1024 + GMAC & 140 & 145 \\
        \hline
        ML-KEM-1024 + CMAC & 143 & 148 \\
        \hline
        ML-KEM-1024 + KMAC & 144 & 149 \\
        \hline
    \end{tabular}
\end{table}

@@

\begin{table}[H]
    \centering
    \caption{UAKE RTT comparison}\label{tbl:uake-rtt}
    \begin{tabular}{|c|c|c|}
        \hline
        KEM & median RTT ($\us$) & average RTT ($\us$) \\
        \hline
        ML-KEM-512 & 145 & 151 \\
        \hline
        ML-KEM-512 + Poly1305 & 103 & 106 \\
        \hline
        ML-KEM-512 + GMAC & 106 & 110 \\
        \hline
        ML-KEM-512 + CMAC & 108 & 112 \\
        \hline
        ML-KEM-512 + KMAC & 109 & 113 \\
        \hline 
        ML-KEM-768 & 215 & 222 \\
        \hline
        ML-KEM-768 + Poly1305 & 144 & 150 \\
        \hline
        ML-KEM-768 + GMAC & 149 & 156 \\
        \hline
        ML-KEM-768 + CMAC & 153 & 160 \\
        \hline
        ML-KEM-768 + KMAC & 154 & 159 \\
        \hline
        ML-KEM-1024 & 310 & 318 \\
        \hline
        ML-KEM-1024 + Poly1305 & 202 & 209 \\
        \hline
        ML-KEM-1024 + GMAC & 212 & 228 \\
        \hline
        ML-KEM-1024 + CMAC & 212 & 218 \\
        \hline
        ML-KEM-1024 + KMAC & 213 & 220 \\
        \hline
    \end{tabular}
\end{table}

@@

\begin{table}[H]
    \centering
    \caption{AKE RTT comparison}\label{tbl:ake-rtt}
    \begin{tabular}{|c|c|c|}
        \hline
        KEM & median RTT ($\us$) & average RTT ($\us$) \\
        \hline
        ML-KEM-512 & 200 & 213 \\
        \hline
        ML-KEM-512 + Poly1305 & 133 & 138 \\
        \hline
        ML-KEM-512 + GMAC & 139 & 143 \\
        \hline
        ML-KEM-512 + CMAC & 143 & 148 \\
        \hline
        ML-KEM-512 + KMAC & 145 & 151 \\
        \hline 
        ML-KEM-768 & 294 & 301 \\
        \hline
        ML-KEM-768 + Poly1305 & 190 & 196 \\
        \hline
        ML-KEM-768 + GMAC & 197 & 210 \\
        \hline
        ML-KEM-768 + CMAC & 202 & 208 \\
        \hline
        ML-KEM-768 + KMAC & 204 & 210 \\
        \hline
        ML-KEM-1024 & 512 & 511 \\
        \hline
        ML-KEM-1024 + Poly1305 & 266 & 273 \\
        \hline
        ML-KEM-1024 + GMAC & 273 & 282 \\
        \hline
        ML-KEM-1024 + CMAC & 280 & 287 \\
        \hline
        ML-KEM-1024 + KMAC & 282 & 288 \\
        \hline
    \end{tabular}
\end{table}

%TODO: Two tables, one-time MAC and many-time MAC, separate encap and decap in 2 tables
%TODO: Look at OpenSSL for GMAC with no AES-256. 
%TODO: Also setup networking experiment on AWS

>>>>

%%%% 1. DOCUMENTCLASS %%%%
%
% TODO: need more thorough references
% TODO: need better intro and conclusion (but later)
% \documentclass[journal=tches,submission,notanonymous]{iacrtrans}
\documentclass[journal=tches,submission]{iacrtrans}
%%%% NOTES:
% - Change "journal=tosc" to "journal=tches" if needed
% - Change "submission" to "final" for final version
% - Add "notanonymous" to reveal authors
% - Add "spthm" for LNCS-like theorems

@@

\subsection{Choosing a message authenticator}
For the $\texttt{ML-KEM}^+$ implementation, we instantiated $\mac$ with a selection that covered a wide range of MAC designs, including Poly1305 \cite{bernstein2005poly1305}, GMAC \cite{mcgrew2004galois}, CMAC \cite{iwata2003omac}\cite{black2000cbc}, and KMAC \cite{kelsey2016sha}.

% TODO: What is Poly1305

% TODO: What is GMAC

% TODO: What is CMAC

% TODO: What is KMAC

% TODO: We benchmarked MAC performance computing tags on various sizes of ML-KEM ciphertexts

@@

% TODO: How is ML-KEM different from Kyber round 3?

@@

    \begin{tabular}[t]{|cc|}
        \hline
        \multicolumn{2}{|c|}{\bf ML-KEM-512} \\
        \hline
        \multicolumn{2}{|c|}{size parameters (bytes)} \\
        pubkey size & 800 \\
        seckey size & 1632 \\
        ciphertext size & 768 \\
        \hline
        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
        Median & 75945 \\
        Average & 76171 \\
        \hline
    \end{tabular}
    \begin{tabular}[t]{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant} 
        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
        \cline{2-5}
        & Median & Average & Median & Average \\
        \hline
        \texttt{ML-KEM-512} & 91467 & 92065 & 121185 & 121650 \\
        \hline
        \texttt{Kyber512} & 97811 & 98090 & 119937 & 120299 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 93157 & 93626 & 33733 & 33908 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ GMAC & 97369 & 97766 & 37725 & 37831 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ CMAC & 99739 & 99959 & 40117 & 39943 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ KMAC & 101009 & 101313 & 40741 & 40916 \\
        \hline
    \end{tabular}\vspace{0.3cm}

    \begin{tabular}[t]{|cc|}
        \hline
        \multicolumn{2}{|c|}{\bf ML-KEM-768} \\
        \hline
        \multicolumn{2}{|c|}{size parameters (bytes)} \\
        pubkey size & 1184 \\
        seckey size & 2400 \\
        ciphertext size & 1088 \\
        \hline
        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
        Median & 129895 \\
        Average & 130650 \\
        \hline
    \end{tabular}
    \begin{tabular}[t]{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant} 
        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
        \cline{2-5}
        & Median & Average & Median & Average \\
        \hline
        \texttt{ML-KEM-768} & 136405 & 147400 & 186445 & 187529 \\
        \hline
        \texttt{Kyber768} & 153061 & 153670 & 182129 & 182755 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 146405 & 146860 & 43315 & 43463 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ GMAC & 149525 & 150128 & 46513 & 46706 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ CMAC & 153139 & 153735 & 49841 & 50074 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ KMAC & 155219 & 155848 & 52415 & 52611 \\
        \hline
    \end{tabular}\vspace{0.3cm}

    \begin{tabular}[t]{|cc|}
        \hline
        \multicolumn{2}{|c|}{\bf ML-KEM-1024} \\
        \hline
        \multicolumn{2}{|c|}{size parameters (bytes)} \\
        pubkey size & 1568 \\
        seckey size & 3168 \\
        ciphertext size & 1568 \\
        \hline
        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
        Median & 194921 \\
        Average & 195465 \\
        \hline
    \end{tabular}
    \begin{tabular}[t]{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant} 
        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
        \cline{2-5}
        & Median & Average & Median & Average \\
        \hline
        \texttt{ML-KEM-1024} & 199185 & 199903 & 246245 & 247320 \\
        \hline
        \texttt{Kyber1024} & 222351 & 223260 & 258231 & 259067 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 205763 & 206499 & 51375 & 51562 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 208805 & 209681 & 54573 & 54780 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 213667 & 214483 & 59175 & 59408 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 216761 & 217468 & 62269 & 62516 \\
        \hline
    \end{tabular}
\end{table}

@@

\begin{table}[H]
    \centering
    \small
    \caption{KE RTT comparison}\label{tbl:ke-rtt}

    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-512} & 800 & 768 & 92 & 97 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 800 & 784 & 70 & 72 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ GMAC & 800 & 784 & 73 & 76 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ CMAC & 800 & 784 & 75 & 79 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ KMAC & 800 & 784 & 76 & 78 \\
        \hline
    \end{tabular}\vspace{0.3cm}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-768} & 1184 & 1088 & 135 & 140 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 1184 & 1104 & 99 & 104 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ GMAC & 1184 & 1104 & 101 & 105 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ CMAC & 1184 & 1104 & 103 & 109 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ KMAC & 1184 & 1104 & 103 & 107 \\
        \hline
    \end{tabular}\vspace{0.3cm}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-1024} & 1568 & 1568 & 193 & 199 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 1568 & 1584 & 138 & 141 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 1568 & 1584 & 140 & 145 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 1568 & 1584 & 143 & 148 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 1568 & 1584 & 144 & 149 \\
        \hline
    \end{tabular}
\end{table}

@@

\begin{table}[H]
    \centering
    \small
    \caption{UAKE RTT comparison}\label{tbl:uake-rtt}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-512} & 1568 & 768 & 145 & 151 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 1584 & 784 & 103 & 106 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ GMAC & 1584 & 784 & 106 & 110 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ CMAC & 1584 & 784 & 108 & 112 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ KMAC & 1584 & 784 & 109 & 113 \\
        \hline
    \end{tabular}\vspace{0.3cm}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-768} & 2272 & 1088 & 215 & 222 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 2288 & 1104 & 144 & 150 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ GMAC & 2288 & 1104 & 149 & 156 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ CMAC & 2288 & 1104 & 153 & 160 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ KMAC & 2288 & 1104 & 154 & 159 \\
        \hline
    \end{tabular}\vspace{0.3cm}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-1024} & 3136 & 1568 & 310 & 318 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 3152 & 1584 & 202 & 209 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 3152 & 1584 & 212 & 228 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 3152 & 1584 & 212 & 218 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 3152 & 1584 & 213 & 220 \\
        \hline
    \end{tabular}
\end{table}

@@

\begin{table}[H]
    \centering
    \small
    \caption{AKE RTT comparison}\label{tbl:ake-rtt}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-512} & 1568 & 1536 & 220 & 213 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 1584 & 1568 & 133 & 138 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ GMAC & 1584 & 1568 & 139 & 143 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ CMAC & 1584 & 1568 & 143 & 148 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ KMAC & 1584 & 1568 & 145 & 151 \\
        \hline
    \end{tabular}\vspace{0.3cm}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-768} & 2272 & 2176 & 294 & 301 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 2288 & 2208 & 190 & 196 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ GMAC & 2288 & 2208 & 197 & 210 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ CMAC & 2288 & 2208 & 202 & 208 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ KMAC & 2288 & 2208 & 204 & 210 \\
        \hline
    \end{tabular}\vspace{0.3cm}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-1024} & 3136 & 3136 & 512 & 511 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 3152 & 3168 & 266 & 273 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 3152 & 3168 & 273 & 282 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 3152 & 3168 & 280 & 287 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 3152 & 3168 & 282 & 288 \\
        \hline
    \end{tabular}
\end{table}

%TODO: Two tables, one-time MAC and many-time MAC, separate encap and decap in 2 tables
%TODO: Look at OpenSSL for GMAC with no AES-256.
